### 2015-04-10 related.tex
edits

Operation-based RDTs have been widely studied in terms of their algorithmic
properties~\cite{SSS,Burckhardt2014}, and several systems utilize this model to
construct distributed data structures~\cite{Cassandra,Bayou,Tango}. These
systems typically propose to implement the datatypes directly over a cluster of
nodes, and only focus on basic eventual consistency. Hence, these systems
implement custom solutions for durability and fault-tolerance. \name realizes
RDTs stronger consistency models on top of off-the-shelf eventually consistent
distributed stores. In this respect, \name is similar to~\cite{BoltOn} where
causal consistency is achieved through a shim layer on Cassandra, which
explicitly tracks and enforces dependencies between updates.
However,~\cite{BoltOn} does not support user-defined RDTs, automatic contract
classification and transactions.

Since eventual consistency alone is insufficient to build correct applications,
several systems~\cite{Bayou,Pileus,RedBlue} propose a lattice of stronger
consistency levels. Similarly, traditional database processing
systems~\cite{Berenson95} and their replicated variants~\cite{BailisHAT}
propose weaker isolation levels for performance. In these systems, the onus is
on the developer to choose the correct consistency(isolation) level for
operations(transactions). \name relieves the developer of this burden, and
instead expects  contracts expressing declarative visibility requirements.

Our contract language and system model is inspired by the axiomatic description
of RDT semantics proposed by~\cite{Burckhardt2014}. While they use axioms for
formal verification of correctness of an RDT implementation, we utilize them as
a means for the user to express the desired consistency guarantees in the
application. Operational semantics of \name (described in the accompanying tech
report~\cite{techrep}) describe an operational manifestation of our system
model. \cite{Burckhardt2015} also presents an operational model of a replicated
data store that is based on the abstract system model presented in
~\cite{Burckhardt2014}. Their claims about the expressivity and practicality of
the system model vindicate ours. However, both differ in the way they aim to
empower the application programmer. \cite{Burckhardt2015} exposes the
operational model to the application programmer, along with primitives, such as
\cf{push}, \cf{pull} and \cf{fetch}, that make it easy to \emph{implement} the
required consistency and isolation guarantees. In contrast, \name completely
abstracts its operational model, and instead relies on contract classification
to \emph{choose} appropriate consistency and isolation levels. Similar
to~\cite{Burckhardt2014} and~\cite{Burckhardt2015}, our system model does not
incorporate real (i.e., wall-clock) time. Hence, our contract language cannot
describe store semantics such as recency or bounded-staleness guarantees
offered by certain stores~\cite{Pileus}.

Several conditions have been proposed to judge whether an operation on a
replicated data object needs coordination or not. ~\cite{Calm} defines
\emph{logical monotonicity} as a sufficient condition for coordination freedom,
and proposes a consistency analysis that marks code regions performing
non-monotonic reasoning (eg: aggregations, such as \cf{COUNT}) as potential
coordination points. ~\cite{IConfluence} and ~\cite{Sieve} define
\emph{invariant confluence} and \emph{invariant safety}, respectively, as
conditions for safely executing an operation without coordination.
~\cite{Sieve} also proposes a program analysis that conservatively marks certain
operations as \emph{blue} (coordination not required), while marking the
remaining as \emph{red} (coordination required). ~\cite{indigo} requires
programmers to declare application semantics, and the desired
application-specific invariants as formulas in first-order logic. It performs
static analysis on these formulas to determine $I$-offender sets - sets of
operations, which, when performed concurrently, result in violation of one or
more of the stated invariants. For each offending set of operation, if
programmer chooses invariant-violation avoidance over violation repair, the
system employs various techniques, such as escrow reservation, to ensure that
the offending set of operations are effectively serialized. Unlike \name, these
works focus on a coarse-grained classification of consistency as eventual or
strong, and do not focus on finer-grained transaction isolation levels. However,
the analyses that some of them propose relieve programmers of the burden to tag
operations with consistency levels. Indeed, we do consider automatic inference
of consistency contracts from application-specific integrity constraints as the
next step for \name.

>>>>

Operation-based RDTs have been widely studied in terms of their algorithmic
properties~\cite{SSS,Burckhardt2014}, and several systems utilize this model to
construct distributed data structures~\cite{Cassandra,Bayou,Tango}. These
systems typically propose to implement the datatypes directly over a cluster of
nodes, and only focus on basic eventual consistency. Hence, these systems
implement custom solutions for durability and fault-tolerance. \name realizes
stronger consistency models for RDTs on top of off-the-shelf eventually consistent
distributed stores. In this respect, \name is similar to~\cite{BoltOn} where
causal consistency is achieved through a shim layer layered on top of Cassandra, which
explicitly tracks and enforces dependencies between updates.
However,~\cite{BoltOn} does not support user-defined RDTs, automatic contract
classification and transactions.

Since eventual consistency alone is insufficient to build correct applications,
several systems~\cite{Bayou,Pileus,RedBlue} propose a lattice of stronger
consistency levels. Similarly, traditional database processing
systems~\cite{Berenson95} and their replicated variants~\cite{BailisHAT}
propose weaker isolation levels for performance. In these systems, the onus is
on the developer to choose the correct consistency(isolation) level for
operations(transactions). \name relieves the developer of this burden, and
instead only requires contracts to express declarative visibility requirements.

Our contract language and system model is inspired by the axiomatic
description of RDT semantics proposed in~\cite{Burckhardt2014}. While
~\cite{Burckhardt2014} uses axioms for formal verification of
an RDT implementation, we utilize them as a means for the user to express
desired consistency guarantees in the application. The operational semantics
of \name (described in the accompanying tech report~\cite{techrep}) describes
an operational manifestation of our system model. \cite{Burckhardt2015} also
presents an operational model of a replicated data store that is based on
the abstract system model presented in ~\cite{Burckhardt2014}. Their claims
about the expressivity and practicality of the system model vindicate
ours. However, both approaches differ in the way they aim to empower the application
programmer. \cite{Burckhardt2015} exposes the operational model to the
programmer, along with primitives, such as \cf{push}, \cf{pull}
and \cf{fetch}, that make it easy to \emph{implement} required
consistency and isolation guarantees. In contrast, \name completely
abstracts its operational model, and instead relies on contract
classification to automatically \emph{choose} appropriate consistency and isolation
levels. Similar to~\cite{Burckhardt2014} and~\cite{Burckhardt2015}, our
system model does not incorporate real (i.e., wall-clock) time. Hence, our
contract language cannot describe store semantics such as recency or
bounded-staleness guarantees offered by certain stores~\cite{Pileus}.

Several conditions have been proposed to judge whether an operation on a
replicated data object needs coordination or not. ~\cite{Calm} defines
\emph{logical monotonicity} as a sufficient condition for coordination
freedom, and proposes a consistency analysis that marks code regions
performing non-monotonic reasoning (eg: aggregations, such as \cf{COUNT}) as
potential coordination points. ~\cite{IConfluence} and ~\cite{Sieve} define
\emph{invariant confluence} and \emph{invariant safety}, respectively, as
conditions for safely executing an operation without coordination.
~\cite{Sieve} also proposes a program analysis that conservatively marks
certain operations as \emph{blue} (coordination not required), while marking
the remaining as \emph{red} (coordination required). ~\cite{indigo} requires
programmers to declare application semantics, and the desired
application-specific invariants as formulas in first-order logic. It
performs static analysis on these formulas to determine $I$-offender sets -
sets of operations, which, when performed concurrently, result in violation
of one or more of the stated invariants. For each offending set of
operations, if the programmer chooses invariant-violation avoidance over
violation repair, the system employs various techniques, such as escrow
reservation, to ensure that the offending set is effectively
serialized. Unlike \name, these approaches focus on a coarse-grained
classification of consistency as eventual or strong, and do not consider
finer-grained transaction isolation levels. However, these analyses do
relieve programmers of the burden of tagging operations with consistency
levels. We consider automatic inference of consistency contracts from
application-specific integrity constraints as an obvious  next step for \name.
